---
title: "Analyze A/B Test Results"
output: html_notebook
---

```{r}
library(dplyr)
library(ggplot2)
library(microbenchmark)
library(readr)
```

## Wrangle

## Gather and Assess

```{r}
tbl <- read_csv("ab_data.csv")
```

```{r}
head(tbl)
```

```{r}
records <- nrow(tbl)
records
```

```{r}
# Number of records with unique users
records_unique_usr <- length(unique(tbl$user_id))
records_unique_usr
```

```{r}
records_non_unique_usr <- records - records_unique_usr
records_non_unique_usr
```

There are 3894 records with non-unique user IDs.

```{r}
# Preview records with non-unique user IDs
tbl %>%
  filter(duplicated(user_id) | duplicated(user_id, fromLast = TRUE)) %>%
  arrange(user_id) %>%
  slice_head(n = 10)
```

The redundant user IDs have mismatched records! For instance, user `630052` is
part of the `treatment` group. As such, this user should have been served the
`new_page`. Yet user `630052` is recorded as having received both. Similarly,
user `630320` is recorded as being part of the `control` group _and_ the
`treatment` group. So in which group is the user? In both instances, user
`630320` is said to have landed on the `old_page`, which should have been served to the `control` group.

```{r}
treat_old <- tbl %>%
  filter(group == "treatment" & landing_page != "new_page")

nrow(treat_old)
```

```{r}
ctrl_new <- tbl %>%
  filter(group == "control" & landing_page != "old_page")

nrow(ctrl_new)
```

```{r}
mismatch <- bind_rows(treat_old, ctrl_new)
mismatch_cnt <- nrow(mismatch)

mismatch_cnt
```

```{r}
records_non_unique_usr - mismatch_cnt
```

All but _one_ of the non-unique user IDs are due to redundant, mismatched
record pairs.

```{r}
# Return non-unique user IDs _not_ due to redundant, mismatched record pairs.
tbl %>%
  anti_join(mismatch) %>%
  filter(duplicated(user_id) | duplicated(user_id, fromLast = TRUE))
```

```{r}
# Test whether there are any missing values in the data frame as a whole.
colSums(is.na(tbl)) > 0
```

## Clean

### Issue 1: There are 3893 redundant user IDs having mismatched records.

The `control` group should be served the `old_page`, and the `new_group` should
be served the `new_page`.

#### Define: Filter for the properly matched results, storing them in a new data frame.

#### Code

```{r}
tbl_2 <- tbl %>%
  filter(
    (group == "control" & landing_page == "old_page") |
    (group == "treatment" & landing_page == "new_page")
  )
```

#### Test

```{r}
tbl_2 %>%
  filter(
    (group == "control" & landing_page != "old_page") |
    (group == "treatment" & landing_page != "new_page")
  )
```

### Issue 2: There is one other non-unique user ID.

Note: The two records have different timestamps. Since the variables of
interest are `group`, `landing_page`, and `conversion`, one of the records can
be dropped.

#### Define: Drop one of the records.

#### Code

```{r}
tbl_2 <- tbl_2 %>%
  distinct(user_id, .keep_all = TRUE)
```

#### Test

```{r}
tbl_2 %>%
  filter(duplicated(user_id) | duplicated(user_id, fromLast = TRUE))
```