---
title: "Analyze A/B Test Results"
output: html_notebook
---

```{r}
library(boot)
library(dplyr)
library(ggplot2)
library(readr)
```

## Wrangle

## Gather and Assess

```{r}
tbl <- read_csv("ab_data.csv")
```

```{r}
head(tbl)
```

```{r}
records <- nrow(tbl)
records
```

```{r}
# Number of records with unique users
records_unique_usr <- length(unique(tbl$user_id))
records_unique_usr
```

```{r}
records_non_unique_usr <- records - records_unique_usr
records_non_unique_usr
```

There are 3894 records with non-unique user IDs.

```{r}
# Preview records with non-unique user IDs
tbl %>%
  filter(duplicated(user_id) | duplicated(user_id, fromLast = TRUE)) %>%
  arrange(user_id) %>%
  slice_head(n = 10)
```

The redundant user IDs have mismatched records! For instance, user `630052` is
part of the `treatment` group. As such, this user should have been served the
`new_page`. Yet user `630052` is recorded as having received both. Similarly,
user `630320` is recorded as being part of the `control` group _and_ the
`treatment` group. So in which group is the user? In both instances, user
`630320` is said to have landed on the `old_page`, which should have been served
to the `control` group.

```{r}
treat_old <- tbl %>%
  filter(group == "treatment" & landing_page != "new_page")

nrow(treat_old)
```

```{r}
ctrl_new <- tbl %>%
  filter(group == "control" & landing_page != "old_page")

nrow(ctrl_new)
```

```{r}
mismatch <- bind_rows(treat_old, ctrl_new)
mismatch_cnt <- nrow(mismatch)

mismatch_cnt
```

```{r}
records_non_unique_usr - mismatch_cnt
```

All but _one_ of the non-unique user IDs are due to redundant, mismatched
record pairs.

```{r}
# Return non-unique user IDs _not_ due to redundant, mismatched record pairs.
tbl %>%
  anti_join(mismatch) %>%
  filter(duplicated(user_id) | duplicated(user_id, fromLast = TRUE))
```

```{r}
# Test whether there are any missing values in the data frame as a whole.
colSums(is.na(tbl)) > 0
```

## Clean

### Issue 1: There are 3893 redundant user IDs having mismatched records.

The `control` group should be served the `old_page`, and the `new_group` should
be served the `new_page`.

#### Define: Filter for the properly matched results, storing them in a new data frame.

#### Code

```{r}
tbl_2 <- tbl %>%
  filter(
    (group == "control" & landing_page == "old_page") |
    (group == "treatment" & landing_page == "new_page")
  )
```

#### Test

```{r}
tbl_2 %>%
  filter(
    (group == "control" & landing_page != "old_page") |
    (group == "treatment" & landing_page != "new_page")
  )
```

### Issue 2: There is one other non-unique user ID.

Note: The two records have different timestamps. Since the variables of
interest are `group`, `landing_page`, and `conversion`, one of the records can
be dropped.

#### Define: Drop one of the records.

#### Code

```{r}
tbl_2 <- tbl_2 %>%
  distinct(user_id, .keep_all = TRUE)
```

#### Test

```{r}
tbl_2 %>%
  filter(duplicated(user_id) | duplicated(user_id, fromLast = TRUE))
```

## Analyze

### Describe: Probability

```{r}
# Overall probability of a user converting
p_pop <- mean(tbl_2$converted)
p_pop
```

```{r}
# Control group's probability of a user converting
tbl_2_ctrl <- tbl_2 %>%
  filter(group == "control")
ctrl_cnvt_prob <- mean(tbl_2_ctrl$converted)

ctrl_cnvt_prob
```

```{r}
# Treatment group's probability of a user converting
tbl_2_treat <- tbl_2 %>%
  filter(group == "treatment")
treat_cnvt_prob <- mean(tbl_2_treat$converted)

treat_cnvt_prob
```

```{r}
# Actual difference between each group's probability of a user converting
obs_diff <- treat_cnvt_prob - ctrl_cnvt_prob
obs_diff
```

<sup>Note: `control` is subtracted from `treatment` to show how exposure to the
new landing page affects the outcome of interest `conversion`: thus, a negative
number indicates a lower probability of conversion, and a positive number
indicates a higher probability.</sup>

In our data sample, we observe that the probability of conversion for the
treatment group is lower--albeit by only 0.0016. But how can we ensure that this
accurately represents the user population as a whole--and isn't just due to
variability and randomness?

### A/B Test

As a starting point, let us assume that the `old_page` is better ($H_0$). To
accept the `new_page` as better ($H_1$), we will require conclusive evidence
with, say, a 95% chance of being right--ie a 5% chance of being wrong ($α$ =
0.05).

$$
H_0: p_{new} - p_{old} \leq 0\\ H_1: p_{new} - p_{old} > 0
$$

#### Bootstrap

We can use the bootstrap technique to measure the level of uncertainty. By
repeatedly resampling observations from the data set with replacement, we
can obtain, say, 1000 estimates for our statistic ie the difference in
conversion rates between treatment and control ($\overline{x}_1 -
\overline{x}_2$). Then, we can approximate the true difference, that is, the
difference for the overall population by taking the mean of the differences,
i.e.,
$$
\overline{B} = \frac{1}{1000} \sum_{r'=1}^{1000}
\,(\overline{x}_1^{\,*r'} -\overline{x}_2^{\,*r'})
$$
along with its standard deviation:
$$
\mathrm{SD}_{\overline{B}} \,(\overline{x}_1 -\overline{x}_2) = \sqrt{\frac{1}{1000 - 1} \sum_{r'=1}^{1000}
\,(\overline{x}_1^{\,*r'} - \overline{x}_2^{\,*r'} - \overline{B})^2}
$$

##### Resample

```{r}
ab_test_bootfun <- function(data, indices) {
  sample_ctrl <- data[indices, ]
  sample_treat <- data[-indices, ]
  # Probability of a user converting for the bootstrap samples
  sample_ctrl_cnvt_prob <- mean(sample_ctrl$converted)
  sample_treat_cnvt_prob <- mean(sample_treat$converted)
  # Calculate the difference between the conversion rates
  p_diffs <- sample_treat_cnvt_prob - sample_ctrl_cnvt_prob
  return(p_diffs)
}
```

```{r}
set.seed(42)
```

```{r}
reps <- boot(data = tbl_2, statistic = ab_test_bootfun, R = 1000)
```

```{r}
# Collect garbage: free unused memory
invisible(gc())
```

```{r}
# Approximate true difference
p_diffs_mean <- mean(reps$t)
p_diffs_mean
```

```{r}
# And standard deviation
p_diffs_mean_sd <- sd(reps$t)
p_diffs_mean_sd
```

```{r}
ci_up_bound <- p_diffs_mean + p_diffs_mean_sd
ci_low_bound <- p_diffs_mean - p_diffs_mean_sd

sprintf("A substantial majority* of sampled differences lie between %f and %f.", ci_low_bound, ci_up_bound)
```

<sup>*68%, assuming a normal distribution</sup>